### 2024-11-20 paper/paper.qmd
initial rproj structure

---
title: "Predicting the fish hunting population"
subtitle: ""
author: 
  - Shreya Sakura Noskor
thanks: "Code and data are available at: [https://github.com/NotSakura/FisheriesData.git](https://github.com/NotSakura/FisheriesData.git)."
date: today
date-format: long
abstract: ""
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(broom)
library(modelsummary)
library(rstanarm)
library(knitr)
```

# Introduction

# Data {#sec-data}

## Overview

## Methodology and Measurement

## Data Visualization {#sec-graphs}

## Outcome and Predictor variables

# Model

## Model Step- Up

### Model justification

## Model Results

# Results

# Discussion

## Weaknesses and next steps

\newpage

\appendix

# Appendix 1

# Appendix 2

# References